\pkg{dataMaid}: your assistant for data cleaning in \proglang{R}

### 2017-01-03 latex/article_vol2.tex
Memory issues

\begin{table}[htbp]
\footnotesize
\bgroup
\def\arraystretch{1.8}% 
\begin{tabular}{L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}}
& \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
\hline
\vspace{0pt} Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
Purpose & Describe some aspect of the variable, e.g. a central value, its dispersion or level of missingness. & Produce a distribution plot. & Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue. \\
Output (required) & A list with entries: \newline \R{\$feature} - a label for the summary value (as character string); \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot. This code should be standalone, i.e. should include the data if necessary. & A list with entries: \newline \R{\$problem} - a logical identifying whether an issue was found; \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
Output (recommended) & A \R{summaryResult} object, i.e. an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format). & \bgroup
\def\arraystretch{1.8}% 
\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} | p{0.45\linewidth}} \vspace*{-1cm}If \R{doEval} is \R{TRUE}: \newline A plot that is opened by the graphic device in \R{R}. & \vspace*{-1cm}If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above. \end{tabular} \egroup  & A \R{checkResult} object, i.e. an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant. \\
Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
\end{tabular}
\egroup
\caption{Reference information for creating new functions to be used
  as part of the \mbox{summarize-,} visualize-, and check steps. The three
  columns correspond to each of the three function types. %\hl{I think the formatting here is still sort of chaotic. Maybe include horizontal lines everywhere? Is it allowed in JSS? It's impossible to read what's in which row right now.}
}
\label{table.functionTypes}
\end{table}

@@

\subsubsection{Economic memory use} 
Another solution, which is especially relevant to Windows users due to
the unfortunate combination of memory control in this operating system
and \proglang{R}.
\hl{And also just R, right? Du har udkommenterert denne kommentar,
men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
det uden en præcision af hvad det går ud på. Hvad tænker du?}
is simply splitting the two steps performed
by \R{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
afterwards. If the \R{rmarkdown} file is very long, as it will
typically be for very large datasets, keeping this file open in memory
wastes precious memory capacities. Therefore, we advice users to
instead split the two steps. This can be done in the following manner:

>>>>

\begin{table}[htbp]
\footnotesize
\bgroup
\def\arraystretch{1.8}% 
\begin{tabular}{L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}}
& \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
\hline
\vspace{0pt} Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
Purpose & Describe some aspect of the variable, e.g. a central value, its dispersion or level of missingness. & Produce a distribution plot. & Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue. \\
Output (required) & A list with entries: \newline \R{\$feature} - a label for the summary value (as character string); \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot. This code should be standalone, i.e. should include the data if necessary. & A list with entries: \newline \R{\$problem} - a logical identifying whether an issue was found; \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
Output (recommended) & A \R{summaryResult} object, i.e. an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format). & \bgroup
\def\arraystretch{1.8}% 
\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} |
                 p{0.45\linewidth}} \vspace*{-1cm}\raggedright If
                 \R{doEval} is \R{TRUE}: \newline A plot that is
                 opened by the graphic device in \R{R}. &
                                                          \vspace*{-1cm}\raggedright If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above. \end{tabular} \egroup  & A \R{checkResult} object, i.e. an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant. \\
Tools available for producing the function & \R{summaryResult()} & --- & \R{messageGenerator()} \newline \R{checkResult()} 
\end{tabular}
\egroup
\caption{Reference information for creating new functions to be used
  as part of the \mbox{summarize-,} visualize-, and check steps. The three
  columns correspond to each of the three function types. %\hl{I think the formatting here is still sort of chaotic. Maybe include horizontal lines everywhere? Is it allowed in JSS? It's impossible to read what's in which row right now.}
}
\label{table.functionTypes}
\end{table}

@@

\subsubsection{Economic memory use} 
Available memory may pose a problem with larger datasets ---
especially on machiens running Windows. The problem is, that older,
32-bit versions of Windows cannot necessary access all physical RAM,
while newer 64-bit versions of Windows cannot access virtual memory,
which makes \proglang{R} unnecessary slow.  Another solution, which is
especially relevant to Windows users due to %the unfortunate
the memory control for this operating system,
%combination of memory control in this operating system and
%\proglang{R},
%\hl{And also just R, right? Du har udkommenterert denne kommentar,
%men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
%det uden en præcision af hvad det går ud på. Hvad tænker du?}
is simply splitting the two steps performed
by \R{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
afterwards. If the \R{rmarkdown} file is very long, as it will
typically be for very large datasets, keeping this file open in memory
wastes precious memory capacities. Therefore, we advice users to
instead split the two steps. This can be done in the following manner:
